---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, 
"SELECT * 
FROM order_products__prior_tbl 
LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl 
LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}
products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?

```{r}
dbGetQuery(sc, 
"
SELECT order_dow
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_dow
ORDER BY n_orders DESC
")
```

* The most popular hour of doing order? 
* The most reordered products?
* The less reorderd product?
* Plot histogram of number of products in an order.

```{r}
dbGetQuery(sc, 
"SELECT order_id
,   COUNT(1) AS n_products
FROM order_products__prior_tbl
GROUP BY order_id
LIMIT 100
")
```

* What is the most popular department? 

```{r}
dbGetQuery(sc, "DESC orders_tbl")
```


## Whisker, magrittr
```{r}
data=list(name="Bartek")
whisker.render("My name is {{name}}.",data)
```
```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
* For given department, give a list of the 10 most popular products (with templating department).

```{r}
data <- list(department_id = 19)



product_times_ordered_txt <-"
SELECT product_id

,   department_id COUNT(1) as pop_dep

FROM products_tbl
WHERE department_id={{department_id}}
GROUP BY pop_dep
ORDER BY pop_dep DESC
LIMIT 10
"
product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
```


## More on orders (Window Functions)

```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order      
FROM orders_tbl
WHERE user_id <= 10 
")

# prints first value of window. Orders by order number by user and takes the order_id of the first element of the window.
```


* Product that is usually ordered in numerous baskets. 

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```



* Could modify the query and add prduct name?
```{r}
"
SELECT product_name
,   pn.product_id
,   AVG(products_number) AS avg_products_number

FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn

LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON pn.product_id = p.product_id

GROUP BY pn.product_id, product_name
ORDER BY avg_products_number DESC
LIMIT 10
" %>%
dbGetQuery(sc, .)
```

* Product that is usually ordered in small baskets.
```{r}
"
SELECT product_name
,   pn.product_id
,   AVG(products_number) AS avg_products_number

FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn

LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON pn.product_id = p.product_id

GROUP BY pn.product_id, product_name
ORDER BY avg_products_number ASC
LIMIT 10
" %>%
dbGetQuery(sc, .)
```


* For 10 first users and all their orders find days since first order.
```{r}

```
